---
title: "Confidence intervals for conditional survival estimates"
author: "Nayef"
date: "8/3/2022"
output: 
   github_document: 
     toc: true
     number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview 
The [Kaplan-Meier curve](https://en.wikipedia.org/wiki/Kaplan%E2%80%93Meier_estimator) gives unconditional survival probability estimates, $\hat{S}(t)$. In many cases, we are interested in how survival probabilities change as a single person/unit progresses over time - these are *conditional* survival probabilities, $\hat{S}(t_2|t_1)$, or the probability of surviving beyond $t_2$, *given that they have already survived up to* $t_1$. These conditional survival estimates can be directly estimated from the KM curve without much difficulty. However, generating confidence intervals around these estimates is a bit more involved. Here, I use the `{condsurv}` R package to demonstrate how this works. 


## References 
1. `condsurv` R package [docs](https://www.emilyzabor.com/condsurv/)
2. `condsurv` R package [github page](https://github.com/zabore/condsurv)
3. 
4. 


# Libraries 
```{r}
# library("remotes")
# install_github("zabore/condsurv")

library("condsurv")
library(dplyr)
library(survival)

```


# Example 

```{r}
lung2 <- 
  mutate(
    lung,
    os_yrs = time / 365.25
  )


myfit <- survfit(Surv(os_yrs, status) ~ 1, data = lung2)

# ?conditional_surv_est
conditional_surv_est(
  basekm = myfit,
  t1 = 0.5, 
  t2 = 1
)


```

