---
title: "Smoothing the KM estimate"
author: "Nayef"
date: "2022-01-06" 
output: 
   github_document: 
     toc: true
     number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview 

Reference: DF Moore, *"Applied Survival Analysis Using R"*, p32. 


# Libraries 
```{r}
library(survival)
library(asaur)
library(survminer)
library(muhaz)
```

# Example 1

## Dataset: Progression-free survival of gastric cancer
```{r}
# ?gastricXelox
time_months <- gastricXelox$timeWeeks * 7/30.25
delta <- gastricXelox$delta  # 1 for death, 0 for censored 

str(gastricXelox)
stopifnot(nrow(gastricXelox) == 48)
```

## KM curve 
```{r}
km1 <- survfit(Surv(time_months, delta) ~ 1, conf.type = "log-log")

km1

ggsurvplot(km1, data = gastricXelox)
# plot(km1, conf.int = T, mark = "|", xlab = "time_months", ylab = "surv prob")

```

## Smoothed hazard functions
- Uses `muhaz::pehaz` to estimate piecewise exponential hazard function from 
right-censored data. 
- `width = 5` means we split the time axis and "bin" into intervals of length 5 
months. 
```{r}
# ?pehaz
# ?muhaz
hazard_1_month_bin <- pehaz(time_months, delta, width = 1)
hazard_5_month_bin <- pehaz(time_months, delta, width = 5)
hazard_smooth <- muhaz(time_months, delta, 
                       bw.smooth = 20, 
                       b.cor = "left")
hazard_smooth_auto <- muhaz(time_months, delta,
                            bw.method = "local")



plot(hazard_1_month_bin)
lines(hazard_5_month_bin, col = "tomato")
lines(hazard_smooth, col = "deepskyblue3", lwd = 2)
lines(hazard_smooth_auto, col = "blue", lwd = 2)

title("Estimated hazard functions using piecewise exponential estimation \nand kernel-based methods")

```

## Deriving smoothed survival estimtate from smoothed hazard function 
```{r}
hazard_values <- hazard_smooth_auto$haz.est
times <- hazard_smooth_auto$est.grid
n_haz <- length(hazard_values)

smooth_surv <- exp(-cumsum(hazard_values[1:n_haz - 1] * diff(times)))

```

## Plotting KM and smoothed survival estimates 
```{r}
plot(km1, conf.int = F, mark = "|", 
     xlab = "time_months", ylab = "surv prob", 
     xlim = c(0, 30))
lines(smooth_surv ~ times[1:(length(times) - 1)], col = "blue", lwd =2)
title("Survival estimates - KM curve and smoothed curve")
```


# Example 2
```{r}

df <- tibble::tribble(
   ~time, ~event,
   810.8,     0L,
   199.7,     1L,
   504.4,     0L,
  400.39,     1L,
   742.9,     0L,
     976,     0L,
   960.7,     0L,
   204.5,     1L,
   597.1,     0L,
   599.6,     1L,
     493,     0L,
   883.5,     0L,
   261.7,     1L,
   568.3,     0L,
   877.5,     0L,
   358.6,     0L,
     819,     0L,
   508.3,     1L,
   287.1,     0L,
   391.2,     1L,
   611.7,     0L,
     709,     0L
  )

str(df)
stopifnot(nrow(df) == 22)

```

## KM curve 
```{r}
km2 <- survfit(Surv(df$time, df$event) ~ 1, conf.type = "log-log")

km2

ggsurvplot(km2, data = df)
# plot(km1, conf.int = T, mark = "|", xlab = "time_months", ylab = "surv prob")

```


## Smoothed hazard functions
```{r}
hazard_smooth_auto2 <- muhaz(df$time, df$event,
                            bw.method = "local")



plot(hazard_smooth_auto2, col = "blue", lwd = 2)
title("Estimated hazard function using kernel-based method")

```

## Deriving smoothed survival estimtate from smoothed hazard function 
```{r}
hazard_values2 <- hazard_smooth_auto2$haz.est
times2 <- hazard_smooth_auto2$est.grid
n_haz2 <- length(hazard_values2)

smooth_surv2 <- exp(-cumsum(hazard_values2[1:n_haz - 1] * diff(times2)))

```

## Plotting KM and smoothed survival estimates 
```{r}
plot(km2, conf.int = F, mark = "|", 
     xlab = "time_months", ylab = "surv prob")
lines(smooth_surv2 ~ times2[1:(length(times2) - 1)], col = "blue", lwd =2)
title("Survival estimates - KM curve and smoothed curve")
```

